Certified Knowledge Compilation \\ with Application to Formally Verified Model Counting

### 2024-01-06 doc/JAIR/jair-paper.tex
fix: typos

We step through portions of the file to provide a better understanding of the CPOG proof framework.
Figure
\ref{fig:eg4:all}(D) shows the defining clauses that are implicitly
defined by the POG operation declarations.  These do not appear in the
CPOG file.  Referring back to the declarations of the sum nodes in
Figure \ref{fig:eg4:all}(C), we can see that the declaration of node
$\nodes_7$ has clause IDs 7 and 10 as the hint.  We can see in Figure
\ref{fig:eg4:all}(A) that these two clauses form a RUP proof for the clause
$\obar{p}_5 \lor \obar{p}_6$, showing that the two children of $\nodes_7$
have disjoint models.  Similarly, node $\nodes_{10}$ is declared as having
clause IDs 16 and 19 as the hint.  These form a RUP proof for the clause
$\obar{p}_8 \lor \obar{p}_9$, showing that the two children of
$\nodes_{10}$ have disjoint models.

@@

Figure \ref{fig:eg4:all}(E) provides the sequence of assertions
leading to unit clause 36, consisting of the literal $s_{10}$.  This clause indicates that $\nodes_{10}$ is implied by the input clauses, i.e.,
any total assignment $\assign$
satisfying the input clauses must have its extension to $\eassign$ yield $\eassign(s_{10}) = 1$.
Working backward, we can see that
clause 35 indicates that variable $s_{10}$ will be implied by the input
clauses when $\assign(x_1) = 0$. Clause 34 indicates that node $p_9$ will
be implied by the input clauses when $\assign(x_1) = 1$, while defining clause 24 shows that node $s_{10}$ will be implied
by the input clauses when $\eassign{p_9} = 1$.  These three clauses
serve as the
hint for clause 36.

@@

Figure \ref{fig:eg4:all}(F) shows the RUP proof steps required to
delete the input clauses.  Consider the first of these, deleting
input clause $\obar{x}_1 \lor x_3 \lor \obar{x}_4$.  The requirement is to show
that there is no total assignment $\assign$ that falsifies this clause but extends to $\eassign$ such that  $\assign(s_{10}) = 1$.
The proof proceeds by first assuming that the clause is false, requiring
$\assign(x_1) = 1$, $\assign(x_3) = 0$, and $\assign(x_4) = 1$.  The hint then consists of unit
clauses (e.g., clause 36 asserting that $\eassign(p_{10}) = 1$) or
clauses that cause unit propagation.  Hint clauses 8 and 10 force the
assignments $\eassign(p_5) = \eassign(p_6) = 0$.  These, plus hint clause 12 force
$\eassign(s_7) = 0$.  This, plus hint clauses 16 and 21 force $\eassign(p_8) = \eassign(p_9) = 0$, leading,
via clause 22, to $\eassign(s_{10}) = 0$.  But this contradicts clause 36,
completing the RUP proof.  The deletion hints for the other input
clauses follow similar patterns---they work from the bottom nodes of
the POG upward, showing that any total assignment that falsifies the clause
must, when extended, have $\eassign(s_{10}) = 0$.

>>>>

We step through portions of the file to provide a better understanding of the CPOG proof framework.
Figure
\ref{fig:eg4:all}(D) shows the defining clauses that are implicitly
defined by the POG operation declarations.  These do not appear in the
CPOG file.  Referring back to the declarations of the sum nodes in
Figure \ref{fig:eg4:all}(C), we can see that the declaration of node
$\nodes_7$ has clause IDs 7 and 10 as the hint.  We can see in Figure
\ref{fig:eg4:all}(D) that these two clauses form a RUP proof for the clause
$\obar{p}_5 \lor \obar{p}_6$, showing that the two children of $\nodes_7$
have disjoint models.  Similarly, node $\nodes_{10}$ is declared as having
clause IDs 16 and 19 as the hint.  These form a RUP proof for the clause
$\obar{p}_8 \lor \obar{p}_9$, showing that the two children of
$\nodes_{10}$ have disjoint models.

@@

Figure \ref{fig:eg4:all}(E) provides the sequence of assertions
leading to unit clause 36, consisting of the literal $s_{10}$.  This clause indicates that $\nodes_{10}$ is implied by the input clauses, i.e.,
any total assignment $\assign$
satisfying the input clauses must have its extension to $\eassign$ yield $\eassign(s_{10}) = 1$.
Working backward, we can see that
clause 35 indicates that variable $s_{10}$ will be implied by the input
clauses when $\assign(x_1) = 0$. Clause 34 indicates that node $p_9$ will
be implied by the input clauses when $\assign(x_1) = 1$, while defining clause 24 shows that node $s_{10}$ will be implied
by the input clauses when $\eassign(p_9) = 1$.  These three clauses
serve as the
hint for clause 36.

@@

Figure \ref{fig:eg4:all}(F) shows the RUP proof steps required to
delete the input clauses.  Consider the first of these, deleting
input clause $\obar{x}_1 \lor x_3 \lor \obar{x}_4$.  The requirement is to show
that there is no total assignment $\assign$ that falsifies this clause but extends to $\eassign$ such that $\eassign(s_{10}) = 1$.
The proof proceeds by first assuming that the clause is false, requiring
$\assign(x_1) = 1$, $\assign(x_3) = 0$, and $\assign(x_4) = 1$.  The hint then consists of unit
clauses (e.g., clause 36 asserting that $\eassign(s_{10}) = 1$) or
clauses that cause unit propagation.  Hint clauses 8 and 10 force the
assignments $\eassign(p_5) = \eassign(p_6) = 0$.  These, plus hint clause 12 force
$\eassign(s_7) = 0$.  This, plus hint clauses 16 and 21 force $\eassign(p_8) = \eassign(p_9) = 0$, leading,
via clause 22, to $\eassign(s_{10}) = 0$.  But this contradicts clause 36,
completing the RUP proof.  The deletion hints for the other input
clauses follow similar patterns---they work from the bottom nodes of
the POG upward, showing that any total assignment that falsifies the clause
must, when extended, have $\eassign(s_{10}) = 0$.
